---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

[![R-CMD-check](https://github.com/desanou/mglasso/actions/workflows/basic.yml/badge.svg)](https://github.com/desanou/mglasso/actions/workflows/basic.yml)

<!-- README.md is generated from README.Rmd. Please edit that file -->

## Method 

This repository provides an implementation of the `MGLasso` (Multiscale Graphical Lasso) algorithm: a novel approach for estimating sparse Gaussian Graphical Models with the addition of a group-fused Lasso penalty.

`MGLasso` is described in the paper [Inference of Multiscale Gaussian Graphical Model](https://desanou.github.io/multiscale_glasso/). `MGLasso` has three major contributions:

-   We simultaneously infer a network and estimate a clustering structure.

-   We use a continuation with Nesterov smoothing in a shrinkage-thresholding algorithm ([`CONESTA`](https://arxiv.org/abs/1605.09658), Hadj-Selem et al. 2018) to solve the optimization problem .

-   We show numerically that `MGLasso` performs better than [`GLasso`](https://arxiv.org/abs/0708.3517)(Friedman et al. 2007) in terms of support recovery.

`MGLasso` package available on [CRAN](https://CRAN.R-project.org/package=mglasso) is based on the python implementation of the solver `CONESTA` available in [pylearn-parsimony](https://github.com/neurospin/pylearn-parsimony) library.


## Package requirements  

-   Install the `reticulate` package.

```{r eval = FALSE}
install.packages('reticulate')
```

-   Check if Python engine is available on the system  

```{r eval = FALSE}
reticulate::py_available()
```

  + If not available   

```{r eval = FALSE}
reticulate::install_miniconda()
```

-   Install `MGLasso`  

```{r eval = FALSE}
install.packages('mglasso')
```

-   Install `MGLasso` python dependencies  

```{r eval = FALSE}
mglasso::install_conesta()
```

An example of use is given below.


## Vignette 

### Installation

```{r eval = FALSE}
remotes::install_github("desanou/mglasso")
library(mglasso)
```

```{r}
library(mglasso)
install_conesta()
```

### Basic Usage

1. Simulate some block diagonal model  

```{r  }
library(Matrix)
n = 50
K = 3
p = 9
rho = 0.85
blocs <- list()

for (j in 1:K) {
  bloc <- matrix(rho, nrow = p/K, ncol = p/K)
  for(i in 1:(p/K)) { bloc[i,i] <- 1 }
  blocs[[j]] <- bloc
}

mat.covariance <- Matrix::bdiag(blocs)
Matrix::image(mat.covariance)
```

  + True cluster partition  

```{r}
rep(1:3, each = 3)
```

  + Simulate gaussian data from the covariance matrix  

```{r}
set.seed(11)
X <- mvtnorm::rmvnorm(n, mean = rep(0,p), sigma = as.matrix(mat.covariance))
```

2. Launch algorithm    

```{r}
X <- scale(X)    
res <- mglasso(X, lambda1 = 0.1, lambda2_start = 0.1, fuse_thresh = 1e-3)
```

3. Plot results compact version   

  + Estimated regression vectors  

```{r fig.width=10, fig.height=10}
plot_mglasso(res)
```

`level9` denotes a partition with `9` clusters. We observe a shrinkage effect in the estimated coefficients due to the fuse-group lasso penalty parameter.

  + Estimated clustering partitions  

```{r  }
res$out$level9$clusters
res$out$level7$clusters
res$out$level4$clusters
res$out$level3$clusters
res$out$level1$clusters
```

The uncovered partition obtained while increasing $\lambda_2$ is a hierarchical partition under some constraints.

## Reference

Edmond, Sanou; Christophe, Ambroise; Geneviève, Robin; (2022): Inference of Multiscale Gaussian Graphical Model. ArXiv. Preprint. <https://doi.org/10.48550/arXiv.2202.05775>
